Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.

### 2024-01-12 apples_and_oranges.qmd
several changes to code

# Step 2: turn dataframes to long
# A: turn long the rows with active
turn_long_active_type <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, descr_active, psy_or_med, baseline_n_active, cohens_d_active) %>% 
  pivot_longer(cols = c(cohens_d_active), 
               names_to = "arm_effect_size", 
               values_to = "cohens_d")

# also rename active_type to treatment for the merge below.
turn_long_active_type <- rename(turn_long_active_type, treatment = descr_active,
                                baseline_n = baseline_n_active)

# # to illustrate the issue, here we have one study with two controls for which the active at the moment, also exists twice. 
# turn_long_active_type[turn_long_active_type$new_study_id == "Stallard, 2012",]
# # but here another one where the same study reasonably contributes two actives, fluoxetine and duloxetine.
# turn_long_active_type[turn_long_active_type$new_study_id == "Atkinson, 2014",]

# we now need to go through each study id and remove duplicates
turn_long_active_type <-
  turn_long_active_type %>% 
  group_by(new_study_id) %>% 
  distinct(treatment, .keep_all = TRUE)

# # Now checking if this worked with the studies that I used to illustrate the problem above. 
# turn_long_active_type[turn_long_active_type$new_study_id == "Stallard, 2012",]
# turn_long_active_type[turn_long_active_type$new_study_id == "Atkinson, 2014",]
# # also check one which is single to make sure it is kept
# turn_long_active_type[turn_long_active_type$new_study_id == "Ackerson, 1998",]

# B: turn long control rows
turn_long_control_type <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, control_type, psy_or_med, baseline_n_control, cohens_d_control) %>% 
  pivot_longer(cols = c(cohens_d_control), 
               names_to = "arm_effect_size", 
               values_to = "cohens_d")

# also rename active_type to treatment for the merge below.
turn_long_control_type <- rename(turn_long_control_type,treatment = control_type,
                                 baseline_n = baseline_n_control)

# # to illustrate the issue, here we have one study with two controls that are reasonable. 
# turn_long_control_type[turn_long_control_type$new_study_id == "Stallard, 2012",]
# # but here another one where the same study reasonably contributes two placebos
# turn_long_control_type[turn_long_control_type$new_study_id == "Atkinson, 2014",]

# we now need to go through each study id and remove duplicates
turn_long_control_type <-
  turn_long_control_type %>% 
  group_by(new_study_id) %>% 
  distinct(treatment, .keep_all = TRUE)

# # Now checking if this worked with the studies that I used to illustrate the problem above. 
# turn_long_control_type[turn_long_control_type$new_study_id == "Stallard, 2012",]
# turn_long_control_type[turn_long_control_type$new_study_id == "Atkinson, 2014",]
# # also check one which is single to make sure it is kept
# turn_long_control_type[turn_long_control_type$new_study_id == "Ackerson, 1998",]

### Now merge the active and control datasets
df_long_for_metan <-rbind(turn_long_active_type, turn_long_control_type)
#dim(df_long_for_metan ) #check dimension
#make sure no study lost
#length(unique(df_long_for_metan$new_study_id ) ) == # #length(unique(df_appl_v_orange$new_study_id ) )

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# library(tidyverse)
# 
# 
# # describe studies
# n_unique_studies <- length(unique(df_long_for_metan$new_study_id))
# 
#   
# 
# df_count_studies <- df_long_for_metan %>% 
#   group_by(psy_or_med, arm_effect_size) %>% 
#   count(is.na(cohens_d))
# 
# 
# df_count_studies <-rename(df_count_studies, is_missing = `is.na(cohens_d)`)
# 
# df_count_studies <-
#   df_count_studies %>% 
#   mutate(condition = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
#                              psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
#                              psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
#                              psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
#   )
# 
#   df_count_studies$condition <- factor(df_count_studies$condition) # turn to factor
#   
#   # relevel so that medication control becomes the reference category for the regression
#   df_count_studies$condition <- relevel(df_count_studies$condition, ref = "medication_control")
# 
# df_count_studies_not_missing <- df_count_studies[df_count_studies$is_missing == FALSE,]
```

@@

for(i in 1: length(list_df_simulated)){
  list_df_simulated[[i]] <- list_df_simulated[[i]] %>% 
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

@@

  # relevel so that medication control becomes the reference category for the regression
  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
}

@@

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cbt_fluox_esc_study <-  aggregate_model_results(list_model_1_cbt_fluox_esc_study, condition)

@@

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_no_wl <-  aggregate_model_results(list_model_1_no_wl, condition)

@@

# apply the metareg function to the list of cbt, fluox, esc
list_model_1_no_wl <- run_metaregression(list_no_wl_or_cau , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_no_wl_or_cau <-  aggregate_model_results(list_model_1_no_wl_or_cau, condition)

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(aggregate_results_overall, caption = "Table 2")
```

@@

z_psy_con_vs_med_con <- 0
for(i in 1: length(list_model_1_no_wl)){
z_psy_con_vs_med_con[i] <- list_model_1_no_wl[[i]]$zval[[4]]
}

@@

update.meta(means_cdrs_baseline_control, 
            subgroup = psy_or_med, 
            tau.common = FALSE)

means_cdrs_baseline_active <- metamean(n = baseline_n_active,
                                        mean = baseline_mean_active,
                                        sd = baseline_sd_active,
                                        studlab = new_study_id,
                                        data = df_with_cdrs,
                                        sm = "MRAW",
                                        fixed = FALSE,
                                        random = TRUE,
                                        method.tau = "REML",
                                        hakn = TRUE,
                                        title = "CDRS Scores at Baseline")
summary(means_cdrs_active)

update.meta(means_cdrs_baseline_active, 
            subgroup = psy_or_med, 
            tau.common = FALSE)

@@

library(meta)
means_hamd_baseline_control <- metamean(n = baseline_n_control,
                                        mean = baseline_mean_control,
                                        sd = baseline_sd_control,
                                        studlab = new_study_id,
                                        data = df_with_hamd,
                                        sm = "MRAW",
                                        fixed = FALSE,
                                        random = TRUE,
                                        method.tau = "REML",
                                        hakn = TRUE,
                                        title = "HAMD Scores at Baseline")
summary(means_hamd_baseline_control)

update.meta(means_hamd_baseline_control, 
            subgroup = psy_or_med, 
            tau.common = FALSE)

means_hamd_baseline_active <- metamean(n = baseline_n_active,
                                       mean = baseline_mean_active,
                                       sd = baseline_sd_active,
                                       studlab = new_study_id,
                                       data = df_with_hamd,
                                       sm = "MRAW",
                                       fixed = FALSE,
                                       random = TRUE,
                                       method.tau = "REML",
                                       hakn = TRUE,
                                       title = "HAMD Scores at Baseline")
summary(means_hamd_baseline_active)

update.meta(means_hamd_baseline_active, 
            subgroup = psy_or_med, 
            tau.common = FALSE)

>>>>

# Step 2: turn dataframes to long
# A: turn long the rows with active
active_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, descr_active, psy_or_med, baseline_n_active, cohens_d_active,
                 baseline_mean_active, baseline_sd_active, post_mean_active, post_sd_active) %>% 
  rename(treatment = descr_active) %>% 
  rename_with(~ str_remove(., "_active")) %>% 
  mutate(arm_effect_size = "cohens_d_active") %>% # chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

control_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, descr_control, psy_or_med, baseline_n_control, cohens_d_control,
                 baseline_mean_control, baseline_sd_control,post_mean_control, post_sd_control) %>% 
  rename(treatment = descr_control) %>% 
  rename_with(~ str_remove(., "_control")) %>% 
  mutate(arm_effect_size = "cohens_d_control") %>% #chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

df_long_for_metan <- rbind(active_df, control_df)

#dim(df_long_for_metan ) #check dimension
#make sure no study lost
#length(unique(df_long_for_metan$new_study_id ) ) == # #length(unique(df_appl_v_orange$new_study_id ) )

@@

@@

list_df_simulated[[i]]$arm_effect_size <- factor(list_df_simulated[[i]]$arm_effect_size)
  list_df_simulated[[i]] <- list_df_simulated[[i]] %>% 
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

@@

  # relevel so that medication control becomes the reference category for the regression
#  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
list_df_simulated[[i]]$four_level_var <-   fct_relevel(list_df_simulated[[i]]$four_level_var,
                                                       "medication_control",
                                                       "medication_active", 
                                                       "psychotherapy_control",
                                                       "psychotherapy_active")
}

@@

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cbt_fluox_esc_study <-  aggregate_model_results(list_model_1_cbt_fluox_esc_study, condition)
aggregate_results_cbt_fluox_esc_study <- cbind(aggregate_results_cbt_fluox_esc_study, condition)

@@

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_no_wl <-  aggregate_model_results(list_model_1_no_wl, condition)
aggregate_results_no_wl <- cbind(aggregate_results_no_wl, condition)

@@

# apply the metareg function to the list of cbt, fluox, esc
list_model_1_no_wl_or_cau <- run_metaregression(list_no_wl_or_cau , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_no_wl_or_cau <-  aggregate_model_results(list_model_1_no_wl_or_cau, condition)
aggregate_results_no_wl_or_cau <- cbind(aggregate_results_no_wl_or_cau, condition)

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(aggregate_results_overall, caption = "Table 2", digits = 2
              )
```

@@

z_psy_con_vs_med_con <- 0
for(i in 1: length(list_model_1_no_wl)){
z_psy_con_vs_med_con[i] <- list_model_1_no_wl[[i]]$zval[condition=="psychotherapy_control"]

@@

subgroup_cdrs_baseline_control <- update.meta(means_cdrs_baseline_control, 
            subgroup = psy_or_med, 
            tau.common = FALSE)

means_cdrs_baseline_active <- metamean(n = baseline_n_active,
                                        mean = baseline_mean_active,
                                        sd = baseline_sd_active,
                                        studlab = new_study_id,
                                        data = df_with_cdrs,
                                        sm = "MRAW",
                                        fixed = FALSE,
                                        random = TRUE,
                                        method.tau = "REML",
                                        hakn = TRUE,
                                        title = "CDRS Scores at Baseline")
#summary(means_cdrs_baseline_active)

subgroup_cdrs_baseline_active <- update.meta(means_cdrs_baseline_active, 
            subgroup = psy_or_med, 
            tau.common = FALSE)

@@

library(meta)
means_hamd_baseline_control <- metamean(n = baseline_n_control,
                                        mean = baseline_mean_control,
                                        sd = baseline_sd_control,
                                        studlab = new_study_id,
                                        data = df_with_hamd,
                                        sm = "MRAW",
                                        fixed = FALSE,
                                        random = TRUE,
                                        method.tau = "REML",
                                        hakn = TRUE,
                                        title = "HAMD Scores at Baseline")
#summary(means_hamd_baseline_control)

subgroup_hamd_baseline_control <- update.meta(means_hamd_baseline_control, 
            subgroup = psy_or_med, 
            tau.common = FALSE)

means_hamd_baseline_active <- metamean(n = baseline_n_active,
                                       mean = baseline_mean_active,
                                       sd = baseline_sd_active,
                                       studlab = new_study_id,
                                       data = df_with_hamd,
                                       sm = "MRAW",
                                       fixed = FALSE,
                                       random = TRUE,
                                       method.tau = "REML",
                                       hakn = TRUE,
                                       title = "HAMD Scores at Baseline")
#summary(means_hamd_baseline_active)

subgroup_hamd_baseline_active <- update.meta(means_hamd_baseline_active, 
            subgroup = psy_or_med, 
            tau.common = FALSE)
